---
title: "Univariate Graphing Assignment"
author: "Pushpak Sunil Rane"
from: markdown+emoji
format:
  pdf:
    highlight-style: a11y
    theme: 
      light: flatly
      dark: darkly
    highlight: pygments
execute: 
  warning: false
  messages: false
---

# Describing Distribution

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load necessary libraries needed.

```{r}
# Load libraries
library(sjPlot); library(ggplot2); library(ggpubr); library(here);library(gtsummary)

# Import data
load(here("data/environmental_impact_clean.Rdata"))
```

## Environmental_impacts_freshwater_use

"environmental_impacts_freshwater_use" variable represents environmental impact of using freshwater (use in km3) on diet and nutrition in certain regions.

```{r}
summary(region$environmental_impacts_freshwater_use)
```

```{r}
IQR(region$environmental_impacts_freshwater_use)
sd(region$environmental_impacts_freshwater_use)
```

Histogram for variable "environmental_impacts_freshwater_use" in km3 units.

```{r}
# this function is from the ggpubr package
gghistogram(data=region, x = "environmental_impacts_freshwater_use", add = "mean", 
            fill = "skyblue", add_density = TRUE, 
            xlab = "Freshwater use (in km3)",
            title = "Environmental Impact of Freshwater Use (in km3)")
```

The histogram titled “Environmental Impact of Freshwater Use (in km³)” depicts the distribution of impact of freshwater use on diet and nutrition in various regions with 351 observations and . The data seems to be right-skewed, indicating more frequent lower values with a long tail of higher values. The mean of environmental impact of freshwater use is around 30.6096 km³, with the median of 2.3623 km³. The interquartile range (IQR) is 10.71601 which spans from 0.4669 km3 to 11.1830 km³, showing where the middle 50% of the data lies that is middle of data from variable. The range of the data is huge with a standard deviation of 133.1164 km³. There seem some outliers that are values outside this range or due to data entry errors.

## Environmental_impacts_land_use

"environmental_impacts_land_use" variable represents environmental impact of land (use in Mkm2) on diet and nutrition in certain regions.

```{r}
summary(region$environmental_impacts_land_use)
```

```{r}
IQR(region$environmental_impacts_land_use, na.rm = TRUE)
sd(region$environmental_impacts_land_use, na.rm = TRUE)
```

Histogram for variable "environmental_impacts_land_use" in Mkm2

```{r}
# this function is from the ggpubr package 
gghistogram(data=region, x = "environmental_impacts_land_use", 
            add = "mean", 
            fill = "green", add_density = TRUE,              
            xlab = "Land use (in Mkm2)", 
            title = "Environmental Impact of Land Use (in Mkm2)")
```

The histogram titled “Environmental Impact of Land Use (in Mkm2)” illustrates how impact of land use on diet and nutrition is distributed in million square kilometers (Mkm²). The data appears to be right-skewed, suggesting that lower values are more common, with a long tail extending towards higher values. The average land use is approximately 1035.65 Mkm², while the median is slightly lower at 57.97 Mkm². The interquartile range (IQR) is 400.3397 which spans from 10.12 Mkm² to 410.46 Mkm², indicating where the middle 50% of the data falls. The overall range of the data is quite large with a standard deviation of 4220.845 Mkm². There are some outliers present, which could be values that fall outside this range or may result from data entry mistakes.

## Region

"region" variable represents regions or places for analyzing impact of it by environmental factors on diet and nutrition.

```{r}
table(region$region,useNA = "always")
```

```{r}
prop.table(table(region$region))
```

```{r}
plot_frq(region$region) + labs(title = "Histogram for Regions", x = "Region", 
                               y = "Frequency") +
  theme(axis.text.x = element_text(angle = 75, vjust = 1, hjust = 1)) 
```

The histogram titled “Histogram for Regions” illustrates how often different regions appear in the dataset. It reveals that almost all regions have same frequency count than Polynesia that has 1 count only. The data for region “Asia” is standing out as the leading region. Every region make 7% of data (i.e. number of observations (N=13)) besides Polynesia which makes 0.3% of data (i.e. number of observations (N=1))

## Disagg_value

*"[disagg_value]{.underline}*" variable represents diet items(fruits,vegetables,beef,etc) in particular region.

```{r}
table(region$disagg_value,useNA = "always")
```

```{r}
prop.table(table(region$disagg_value))
```

```{r}
plot_frq(region$disagg_value) + labs(title = "Histogram for Disagg_value", x = "Disagg_value", y = "Frequency")
```

The plot illustrates how often different disagg_value (i.e. diet) appear in the dataset. It reveals that almost all disagg_value have same frequency count than total which contains all other disagg_values that are there in that region having 27 which is higher than other. The data for region “total” is standing out as the leading value. Every value makes 7.7% of data (i.e. number of observations (N=26)) besides total which makes 8.0% of data (i.e. number of observations (N=27)).
